---
title: "Introduction to ggbiome"
author: "Guillem Bagaria"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ggbiome}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`ggbiome` provides functions to obtain mean annual temperature,
mean annual precipitation and Whittaker's biome type from geographical
coordinates, and to plot locations over the biome diagram.


## Get data functions

These functions are intended to obtain data and create objects to be plotted.
They involve creating a spatial object of Whittaker's biomes, and downloading 
climatic data and obtaining the biome using geographical coordinates.

### `gd_get_biomes_spdf`

It returns a SpatialPolygonsDataFrame object of the Whittaker's biomes
modified by Ricklefs (2008) in function of mean annual temperature and mean 
annual precipitation.

### `gd_get_biome`

This function takes a data frame with geographical coordinates or climatic data 
and returns the same data frame with extra columns of climatic data (if not 
provided) and biome. It also accepts single values for both latitude and
longitude, or both mean annual temperature and mean annual precipitation.
Climatic data is obtained from WorldClim 1.4 using the `RFc` package.


## Visualization functions

The aim of these functions is to produce biome diagrams. They plot either
a diagram of the spatial object of the Whittaker's biomes, or site locations 
over the same diagram.

### `vis_biome`

This function produces a ggplot object of the Whittaker's biomes from a
SpatialPolygonsDataFrame object obtained with `gd_get_biomes_spdf`.
It shows each biome as a colored area according to mean annual temperature
and mean annual precipitation.

### `vis_location_biome`

It produces a ggplot object showing the Whittaker's biomes as colored areas 
according to mean annual temperature and mean annual precipitation using the
function `vis_biome`, and adds sites on it according to their climatic
values, obtained using the function `gd_get_biome` if not provided.


## Usage

### Obtaining climatic and biome type data

`gd_get_biome` expects either a data frame with site coordinates or climatic 
data (indicating column names for columns that contain this data), or values
of both latitude (`si_lat`) and longitude (`si_long`) in decimal degrees,
or both mean annual temperature (`si_mat`) in degrees Celsiu and mean annual
precipitation (`si_map`) in millimeters. It returns a data frame with the 
original data and extra columns of mean annual temperature, mean annual 
precipitation (if not provided) and biome:

```{r obtain-data}
# Load package
library(ggbiome)

# Create a data frame with coordinates
sites <- data.frame(site = c("A", "B"), lat = c(69.49, 41.43),
                    long = c(27.23, 2.07))

# Get temperature, precipitation and biomes
gd_get_biome(sites, si_lat = "lat", si_long = "long")

# Alternatively, for one site
gd_get_biome(si_mat = 15.8, si_map = 622)
```


### Plotting locations over the biome diagram

`vis_location_biome` expects the same input as `gd_get_biome`, and returns a
ggplot object showing the biomes with sites over it: 

```{r plot-data, fig.width=7, fig.height=4}
# Create a ggplot diagram of the biomes with sites
vis_location_biome(sites, si_lat = "lat", si_long = "long",
                   point_labels = sites$site)

# Alternatively, for one site
vis_location_biome(si_mat = 15.8, si_map = 622, col_fill = 'red')
```
